The Noisy Work of Uncertainty Visualisation Research: A Review

### 2024-10-07 paper.qmd
checked libraries were all doing something

# load Libraries
library(tidyverse)
library(biscale)
library(RColorBrewer)
library(scales)
library(sf)
library(ggrepel)
library(urbnmapr)
library(flextable)
```

@@

# Picture 1
p1 <- ib_data %>%
  summarise(avg = mean(x),
          conf_95a = quantile(x, probs=c(0.025)),
          conf_95b = quantile(x, probs=c(0.975))) %>%
  ggplot(aes(y="NA")) +
  geom_point(aes(x=avg)) +
  geom_errorbar(aes(xmin = conf_95a, xmax = conf_95b), width = 0.1) +
  scale_x_continuous(name = "INCHES OF SNOW",
                     breaks=seq(0,19),
                     labels= ggplot2:::interleave(as.character(c(seq(0,18, 2), 19)), rep("", 11))[c(0:19, 21)],
                     limits=c(0,19)) +
  theme_classic() +
  theme(axis.line.y=element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        aspect.ratio=1/10)

@@

# Picture 3
# library(ggpattern)
p3 <- ib_data |>
  mutate(x = ifelse(x>18, 18, x),
         binx = cut(x, 
                     breaks=seq(0,18,2), 
                     labels= paste0(seq(0,16,2), sep = "-", seq(2,18,2)))) |>
  group_by(binx) |>
  summarise(n = n()) |>
  mutate(Probability = n / sum(n),
         csum = rev(cumsum(rev(Probability))), 
         pos = Probability/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Probability/2, pos)) |>
  ggplot(aes(x="", y=Probability, fill=binx)) +
  geom_bar(stat="identity", width=1) +
  geom_text_repel(aes(y = pos, label = paste0(round(Probability*100), sep="", "%")),
                   size = 3, nudge_x = 0.6, show.legend = FALSE, segment.color = 'transparent') +
  #geom_label(aes(label = paste0(round(Probability*100), sep="", "%")),
  #          position = position_stack(vjust = 0.5)) + 
  scale_fill_grey() +
  coord_polar("y", start=0) +
  labs(fill = "INCHES OF SNOW") + 
  theme_void() + 
  theme(aspect.ratio=1)

@@

# Picture 8
p8 <- ib_data %>%
  reframe(x = quantile(x, probs=c(0.25, 0.50, 0.75)))|>
  add_row(x = c(0,20)) |>
  arrange(x) |>
  mutate(quantile = c("min", "q1", "med", "q3", "max")) |>
  pivot_wider(names_from = quantile, values_from = x) |>
  ggplot(aes(y="")) +
  #geom_point(aes(x=med)) +
  geom_errorbar(aes(y="", xmin = min, xmax = max), width = 0.2) +
  geom_crossbar(aes(y="", x=med, xmin = q1, xmax = q3), width = 0.5) +
  scale_x_continuous(name = "INCHES OF SNOW",
                     breaks=seq(0,20),
                     labels= ggplot2:::interleave(as.character(c(seq(0,20, 2))), rep("", 11))[1:21],
                     limits=c(0,20)) +
  theme_classic() +
  theme(axis.line.y=element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        aspect.ratio=1/10)

@@

# Get map data
my_map_data <- get_urbn_map("counties", sf = TRUE) |>
  dplyr::filter(state_name=="Iowa")
centroids <- as_tibble(rgeos::gCentroid(as(my_map_data$geometry, "Spatial"), byid = TRUE))
my_map_data$cent_long <- centroids$x
my_map_data$cent_lat <- centroids$y
n <- dim(my_map_data)[1]

# Make palettes
longpal <- rev(colorspace::sequential_hcl(13, palette = "YlOrRd"))
basecols <- longpal[3:10]
breaks <- 21:29 
breakslong <- 18:32 
names(basecols) <- seq(8)
names(longpal) <- -1:11

my_map_data <- my_map_data |>
  dplyr::mutate(temp = 29 - 3*abs(scale(cent_lat) - sin(2*(scale(cent_long)))[,1]), # trend
         highvar = runif(n, min=2, max=4), # high variance
         lowvar = runif(n, min=0, max=2), # low variance
         count_id = row_number()) |>
  pivot_longer(cols=highvar:lowvar, names_to = "variance_class", values_to = "variance") |>
  # add bivariate classes to data
  mutate(bitemp = cut(temp, breaks=breaks, labels=seq(8)),
         bivar = cut(variance, breaks=0:4, labels=seq(4)),
         biclass = paste(bitemp, bivar, sep="-"))|>
  mutate(highlight = ifelse(count_id <= 5, TRUE, FALSE))

@@

# Choropleth Map
p1a <- my_map_data |>
  filter(variance_class=="lowvar") %>%
  ggplot() +
  geom_sf(aes(fill = bitemp, 
              geometry = geometry), colour=NA) + 
  scale_fill_manual(values = basecols) +
  #scale_fill_gradientn(colours = basecols, 
  #                     values=breaks/limits[2],
  #                     limits=limits) +
  theme_void() + 
  theme(legend.position = "none")

@@

# Exceed Prob Map
p2a <- exeed_data |>
  filter(variance_class=="lowvar") %>%
  ggplot() +
  geom_sf(aes(fill = xprob, 
              geometry = geometry), colour=NA) + 
  scale_fill_manual(values = basecols) +
  theme_void() + 
  theme(legend.position = "none")

@@

# Bivariate maps
p2a <- my_map_data |>
  filter(variance_class=="lowvar") %>%
  ggplot() +
  geom_sf(aes(fill = biclass, geometry = geometry), colour=NA) + 
  scale_fill_manual(values = bivariatepal) +
  theme_void() + 
  theme(legend.position = "none")

@@

# VSUP maps
p3a <- my_map_data |>
  filter(variance_class=="lowvar") %>%
  ggplot() +
  geom_sf(aes(fill = biclass, geometry = geometry), colour=NA) + 
  scale_fill_manual(values = VSUP) +
  theme_void() + 
  theme(legend.position = "none")

@@

# Pixel Map
library(Vizumap)

# Low variance map
my_map_data_a <- my_map_data %>%
  filter(variance_class == "lowvar") %>%
  mutate(my_id = seq(n),
         error = variance)

@@

pixel_1a <- my_map_data_a %>%
  as.data.frame() %>%
  select(my_id, bitemp, error) %>%
  read.uv(estimate="bitemp", error="error")

pixel_2a <- my_map_data_a %>% as("Spatial")

@@

# High variance
my_map_data_b <- my_map_data %>%
  filter(variance_class == "highvar") %>%
  mutate(my_id = seq(n),
         error = variance)

@@

pixel_1b <- my_map_data_b %>%
  as.data.frame() %>%
  select(my_id, bitemp, error) %>%
  read.uv(estimate="bitemp", error="error")

pixel_2b <- my_map_data_b %>% 
  as("Spatial")

>>>>

# load Libraries
library(tidyverse)
library(Vizumap)
library(ggplot2)
library(RColorBrewer)
library(scales)
library(sf)
library(ggrepel)
library(urbnmapr)
library(flextable)
library(colorspace)
library(rgeos)
```

@@

# Picture 1
p1 <- ib_data |>
  summarise(avg = mean(x),
          conf_95a = quantile(x, probs=c(0.025)),
          conf_95b = quantile(x, probs=c(0.975))) |>
  ggplot(aes(y="NA")) +
  geom_point(aes(x=avg)) +
  geom_errorbar(aes(xmin = conf_95a, xmax = conf_95b), width = 0.1) +
  scale_x_continuous(name = "INCHES OF SNOW",
                     breaks=seq(0,19),
                     labels= ggplot2:::interleave(as.character(c(seq(0,18, 2), 19)), rep("", 11))[c(0:19, 21)],
                     limits=c(0,19)) +
  theme_classic() +
  theme(axis.line.y=element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        aspect.ratio=1/10)

@@

# Picture 3
p3 <- ib_data |>
  mutate(x = ifelse(x>18, 18, x),
         binx = cut(x, 
                     breaks=seq(0,18,2), 
                     labels= paste0(seq(0,16,2), sep = "-", seq(2,18,2)))) |>
  group_by(binx) |>
  summarise(n = n()) |>
  mutate(Probability = n / sum(n),
         csum = rev(cumsum(rev(Probability))), 
         pos = Probability/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Probability/2, pos)) |>
  ggplot(aes(x="", y=Probability, fill=binx)) +
  geom_bar(stat="identity", width=1) +
  geom_text_repel(aes(y = pos, label = paste0(round(Probability*100), sep="", "%")),
                   size = 3, nudge_x = 0.6, show.legend = FALSE, segment.color = 'transparent') +
  #geom_label(aes(label = paste0(round(Probability*100), sep="", "%")),
  #          position = position_stack(vjust = 0.5)) + 
  scale_fill_grey() +
  coord_polar("y", start=0) +
  labs(fill = "INCHES OF SNOW") + 
  theme_void() + 
  theme(aspect.ratio=1)

@@

# Picture 8
p8 <- ib_data |>
  reframe(x = quantile(x, probs=c(0.25, 0.50, 0.75)))|>
  add_row(x = c(0,20)) |>
  arrange(x) |>
  mutate(quantile = c("min", "q1", "med", "q3", "max")) |>
  pivot_wider(names_from = quantile, values_from = x) |>
  ggplot(aes(y="")) +
  #geom_point(aes(x=med)) +
  geom_errorbar(aes(y="", xmin = min, xmax = max), width = 0.2) +
  geom_crossbar(aes(y="", x=med, xmin = q1, xmax = q3), width = 0.5) +
  scale_x_continuous(name = "INCHES OF SNOW",
                     breaks=seq(0,20),
                     labels= ggplot2:::interleave(as.character(c(seq(0,20, 2))), rep("", 11))[1:21],
                     limits=c(0,20)) +
  theme_classic() +
  theme(axis.line.y=element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        aspect.ratio=1/10)

@@

# Get map data
my_map_data <- get_urbn_map("counties", sf = TRUE) |>
  filter(state_name=="Iowa")
centroids <- as_tibble(gCentroid(as(my_map_data$geometry, "Spatial"), byid = TRUE))
my_map_data$cent_long <- centroids$x
my_map_data$cent_lat <- centroids$y
n <- dim(my_map_data)[1]

# Make palettes
longpal <- rev(sequential_hcl(13, palette = "YlOrRd"))
basecols <- longpal[3:10]
breaks <- 21:29 
breakslong <- 18:32 
names(basecols) <- seq(8)
names(longpal) <- -1:11

my_map_data <- my_map_data |>
  mutate(temp = 29 - 3*abs(scale(cent_lat) - sin(2*(scale(cent_long)))[,1]), # trend
         highvar = runif(n, min=2, max=4), # high variance
         lowvar = runif(n, min=0, max=2), # low variance
         count_id = row_number()) |>
  pivot_longer(cols=highvar:lowvar, names_to = "variance_class", values_to = "variance") |>
  # add bivariate classes to data
  mutate(bitemp = cut(temp, breaks=breaks, labels=seq(8)),
         bivar = cut(variance, breaks=0:4, labels=seq(4)),
         biclass = paste(bitemp, bivar, sep="-"))|>
  mutate(highlight = ifelse(count_id <= 5, TRUE, FALSE))

@@

# Choropleth Map
p1a <- my_map_data |>
  filter(variance_class=="lowvar") |>
  ggplot() +
  geom_sf(aes(fill = bitemp, 
              geometry = geometry), colour=NA) + 
  scale_fill_manual(values = basecols) +
  #scale_fill_gradientn(colours = basecols, 
  #                     values=breaks/limits[2],
  #                     limits=limits) +
  theme_void() + 
  theme(legend.position = "none")

@@

# Exceed Prob Map
p2a <- exeed_data |>
  filter(variance_class=="lowvar") |>
  ggplot() +
  geom_sf(aes(fill = xprob, 
              geometry = geometry), colour=NA) + 
  scale_fill_manual(values = basecols) +
  theme_void() + 
  theme(legend.position = "none")

@@

# Bivariate maps
p2a <- my_map_data |>
  filter(variance_class=="lowvar") |>
  ggplot() +
  geom_sf(aes(fill = biclass, geometry = geometry), colour=NA) + 
  scale_fill_manual(values = bivariatepal) +
  theme_void() + 
  theme(legend.position = "none")

@@

# VSUP maps
p3a <- my_map_data |>
  filter(variance_class=="lowvar") |>
  ggplot() +
  geom_sf(aes(fill = biclass, geometry = geometry), colour=NA) + 
  scale_fill_manual(values = VSUP) +
  theme_void() + 
  theme(legend.position = "none")

@@

# Pixel Map

# Low variance map
my_map_data_a <- my_map_data |>
  filter(variance_class == "lowvar") |>
  mutate(my_id = seq(n),
         error = variance)

@@

pixel_1a <- my_map_data_a |>
  as.data.frame() |>
  select(my_id, bitemp, error) |>
  read.uv(estimate="bitemp", error="error")

pixel_2a <- my_map_data_a |> as("Spatial")

@@

# High variance
my_map_data_b <- my_map_data |>
  filter(variance_class == "highvar") |>
  mutate(my_id = seq(n),
         error = variance)

@@

pixel_1b <- my_map_data_b |>
  as.data.frame() |>
  select(my_id, bitemp, error) |>
  read.uv(estimate="bitemp", error="error")

pixel_2b <- my_map_data_b |> 
  as("Spatial")
